Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-occurrence Where the Hudsons's Bay Lowlands meets the Ontario Shield

### 2024-04-29 analysis/documents/1-data-preparation.qmd
updates

exif <- readr::read_csv(here::here("data", "tabular", "all_exif.csv")) |>
  janitor::clean_names() |>
  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) |>
  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage_2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage/")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = strsplit(image_id, "_") |> 
                  purrr::map(~paste0(.x[2:length(.x)], collapse = "_")) |> 
                  unlist()) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/")))

@@

@@

@@

@@

table_cat <- lu_dat |> 
  dplyr::select(category_code, label) |> 
  dplyr::filter(category_code != "ND") |> 
  dplyr::arrange(category_code) |> 
  gt::gt() |>
  gt::tab_header(
    title = "Table 4: Land cover classes abbreviations corresponding to the Ontatio Far North Land cover dataset.",
    subtitle = "Adapted from the ontario Far North land cover dataset"
  ) |>
  gt::cols_label(
    label = "Land cover class",
    category_code = "Abbreviations",
  ) |> 
  gt::opt_stylize(style = 1, color = "gray")

@@

>>>>

exif <- readr::read_csv(here::here("data", "tabular", "all_exif.csv")) |>
  janitor::clean_names() |>
  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) |>
  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage_2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage/")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = strsplit(image_id, "_") |> 
                  purrr::map(~paste0(.x[2:length(.x)], collapse = "_")) |> 
                  unlist()) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/"))) #|> 
#   dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
#                                                tz = "Canada/Eastern"),
#                 year = lubridate::year(date_time),
#                 month = lubridate::month(date_time, label = TRUE),
#                 day = lubridate::day(date_time),
#                 ymd = paste(year, month, day, sep = "_"))
# 
# exif[
#   exif$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
# exif[
#   exif$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"

@@

anns_wide_animals_joined_mut[
  anns_wide_animals_joined_mut$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
anns_wide_animals_joined_mut[
  anns_wide_animals_joined_mut$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"

@@

camera_days[camera_days$camera_first_day == "2024-05-23 13:31:46",]$camera_first_day <- as.Date("2022-05-23 13:31:46")

@@

all_table <- dplyr::bind_rows(extr) |> 
  dplyr::group_by(value) |> 
  dplyr::summarise(total = sum(coverage_fraction)) |> 
  dplyr::mutate(value = as.factor(value), 
                Percentage = round(total/sum(total)*100, 2))

@@

table_cat <- lu_dat |> 
  dplyr::left_join(all_table, by = dplyr::join_by(cats == value)) |> 
  dplyr::mutate(Percentage = ifelse(is.na(Percentage), 0, Percentage)) |>  
  dplyr::select(category_code, label, Percentage) |> 
  dplyr::filter(category_code != "ND") |> 
  dplyr::arrange(category_code) |> 
  gt::gt() |> 
  gt::cols_label(
    label = "Land cover class",
    category_code = "Abbreviations",
    Percentage = "% Cover in 1km buffers"
  ) |> 
  gt::opt_stylize(style = 1, color = "gray")

@@

#### Process WildTrax data

```{r}
# debugonce(wt_auth)
wildRtrax::wt_auth(force = T)

# debugonce(wt_get_download_summary)
my_projects <- wildRtrax::wt_get_download_summary(sensor_id = 'CAM')

# debugonce(wt_download_report)
my_report <- wildRtrax::wt_download_report(project_id = 2431, sensor_id = 'CAM', reports = "image_report", weather_cols = F) |>
  tibble::as_tibble()
```

### 2024-04-29 analysis/documents/2-data-exploration.qmd
updates

# Fixing error 
# TODO make it more explicit
# image_data[image_data$date_time==max(image_data$date_time, na.rm = T),] |> View()
image_data[image_data$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"

@@

```{r, fig.height=40, fig.width=8}
cam_dep_data_with_status |> 
  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(camera_id, 
                                                              date_deployed,
                                                              decreasing = TRUE))) +
  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = camera_id,
                                     color = status), alpha = 0.15, linewidth = 3) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id,
                                   color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id,
                                   color = status)) +

@@

>>>>

# Fixing error 
# TODO make it more explicit
# image_data[image_data$date_time==max(image_data$date_time, na.rm = T),] |> View()
# image_data[image_data$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
# image_data[image_data$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"

@@

```{r, eval: FALSE fig.height=35, fig.width=7}
#| fig.width: 8
#| fig.height: 35

cam_dep_data_with_status <- cam_dep_data |>
  dplyr::mutate(status = ifelse(date_deployed < date_retrieved, "valid", "invalid")) |>
  dplyr::mutate(status = ifelse(is.na(status), "invalid", status))

cam_dep_data_with_status |>
  ggplot2::ggplot(ggplot2::aes(x = date_deployed, y = reorder(camera_id,
                                                              date_deployed,
                                                              decreasing = TRUE))) +
  ggplot2::geom_segment(ggplot2::aes(xend = date_retrieved, yend = camera_id,
                                     color = status), alpha = 0.15, linewidth = 3) +
  ggplot2::geom_point(ggplot2::aes(x = date_deployed, y = camera_id,
                                   color = status)) +
  ggplot2::geom_point(ggplot2::aes(x = date_retrieved, y = camera_id,
                                   color = status)) +

@@

#### Compare platform

```{r}
wt <- readRDS("../../my_report.rds")
```

```{r}
image_data |> dplyr::filter(plot_id == "P045") |> dplyr::filter(grepl(.data$file_path, pattern = "TrailCamStorage_2/", fixed = T)) |> dplyr::select(image_id, species) |> unique() |> dplyr::pull(species) |> table() |> sort()

wt |> dplyr::filter(grepl(.data$location, pattern = "TC2_P045", fixed = T)) |> dplyr::select(image_id, species_common_name) |> unique() |> dplyr::pull(species_common_name) |> table() |> sort()
```

```{r}
image_data |> dplyr::filter(plot_id == "P045") |> dplyr::filter(grepl(.data$file_path, pattern = "TrailCamStorage/", fixed = T)) |> dplyr::select(image_id, species) |> unique() |> dplyr::pull(species) |> table() |> sort()

wt |> dplyr::filter(grepl(.data$location, pattern = "TC1_P045", fixed = T)) |> dplyr::select(image_id, species_common_name) |> unique() |> dplyr::pull(species_common_name) |> table() |> sort()
```

```{r}
image_data |> dplyr::filter(plot_id == "P080") |> dplyr::filter(grepl(.data$file_path, pattern = "TrailCamStorage/", fixed = T)) |> dplyr::select(image_id, species) |> unique() |> dplyr::pull(species) |> table() |> sort()

wt |> dplyr::filter(grepl(.data$location, pattern = "TC1_P080", fixed = T)) |> dplyr::select(image_id, species_common_name) |> unique() |> dplyr::pull(species_common_name) |> table() |> sort()
```

#### AI results explore

```{r}
library(tidyverse)

labels <- read.csv("../../3_Classifiers/1_species_classifier/labels.csv") |> 
  pivot_longer(OtherBird:Bear, names_to = "Category_true", values_to = "val") |> 
  filter(val != 0) |> select(-val)
preds <- read.csv("../../3_Classifiers/1_species_classifier/predictions.csv") |> 
  pivot_longer(!X, names_to = "Category_pred", values_to = "conf") |> 
  group_by(X) |> 
  arrange(desc(conf)) |> 
  slice(1) |> ungroup()

joined <- labels |> 
  left_join(preds, by=c("id"="X")) |> 
  drop_na() |> 
  mutate(correct = Category_true == Category_pred)

errors <- joined |> filter(!correct)
```

```{r}
ex <- errors |> filter(Category_true == "Moose", Category_pred == "Caribou") |> 
  pull(filepath) |> 
  str_replace("/media/vlucet/TrailCamST/", "/media/vlucet/TrailCamST1/_old/")

for (thefile in ex) {
  # print(file.path("/media/vlucet/TrailCamST1/_old/errors_ex", basename(thefile)))
  file.copy(thefile, file.path("/media/vlucet/TrailCamST1/_old/errors_ex/", basename(thefile)))
}
```

### 2024-04-29 reports/GnC_report/paper/paper.qmd
updates

TODO: update on the tagging progress on WildTrax at the end of April

@@

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units = "days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) |> 
  dplyr::filter(diff_time < 200) # For mustelid error currentlycams_sf_coords TODO

>>>>

About 97% of all images previously tagged in LabelStudio have been uploaded to the WildTrax platform (3,192,316 images), in addition to all the images from the cameras most recently retrieved. The remaining images are still being uploaded. These images are being re-tagged at the species level, and are in various stages of progress, with about half of those folders fully tagged and verified, illustrating the superior speed of tagging on the platform. A more detailed report of this progress and tagging method will be present in the final version of this report, once the revision of the ragging workflow is finalized.

@@

  dplyr::mutate(diff_time = difftime(first_day, camera_first_day, units = "days"),
                ECOZONE_NA = as.factor(ECOZONE_NA)) |> 
  dplyr::filter(diff_time < 200) # For mustelid error currently TODO
